### 2014-06-19 paper/discussion.tex
UV luminosity - SFR

\section{Discussion}

\begin{figure}
\epsscale{1.00}
\plotone{fig/cosmic_variance.pdf}
\caption{Cosmic Variance: The Luminosity Function is made using the DMH catalog 
from the full box and the set of parameters from the small boxes.}
\label{graph_cosmic_variance}
\end{figure}

Q: Mpc/h ?