### 2014-06-19 paper/draft.tex

\documentclass[manuscript]{aastex}
%\documentclass[preprint2]{aastex}
%\usepackage{graphicx}
%\usepackage{graphics}
%\usepackage[dvips]{epsfig}
\usepackage{epstopdf}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{skywalker@galaxy.far.far.away}

\shorttitle{A Connection Between Star Formation Rate and Dark Matter Halos at Z 
\~ 6}
\shortauthors{G\'omez-Cort\'es}

\begin{document}

\title{A Connection Between Star Formation Rate and Dark Matter Halos at 
$Z\sim6$ In 2013 Planck Cosmology.}

\author{F.L. G\'omez-Cort\'es\altaffilmark{1} }
\affil{Departamento de Física, Universidad de los Andes, Colombia}

\begin{abstract}
This work relates baryonic matter and dark matter at redshift $z=5.9$ using 
observational data from CFHTLS \citep{willott13}, HUDF09 
\citep{bouwens06,bouwens12}, UKIDSS and SDXS \citep{mclure09}, and results of 
the Multidark Simulation \citep{riebe13} in a cubic box of $1000 \textrm{Mpc 
h}^{-1}$ length with 2013 Planck Cosmology. The Luminosity Function (LF) is 
fitted via four parameters with the Markov Chain Monte Carlo method. The 
relationship between the Dark Matter Halos Mass and Star Formation Rate is 
obtained using the relationship between the UV  continuum (from the fitted LF) 
and Star Formation Rate (SFR) by \cite{kennicutt98}.Cosmic variance effects are 
studied on smaller boxes of $250 \textrm{Mpc h}^{-1}$ length.

%Luminosity Function - Dark Matter Halos, Star Formation Rate - Dark Matter 
Halos.
\end{abstract}
\keywords{Dark Matter, LF, SFR, High Redshift Galaxies, Reionization}

\include{introduction}

\include{star_formation_rate}

\include{fitting_model}

\include{observations}

\include{discussion}

\section{Summary}

\acknowledgments
%Gracias Totales
%
\appendix

\section{Appendix material}
DEXTER \url{http://dc.zah.uni-heidelberg.de/dexter/ui/ui/custom}

\include{the_bibliography}

\end{document}

%%
%% End of file `sample.tex'.