### 2014-06-20 paper/fitting_model.tex
SFR, Fitting Model, The_Bibliography

\section{The Fitting Model}
The model links dark matter and U-luminosity by a four parameter function.
For each galaxy we have the luminosity:
  \begin{equation}
  L = L_{0} M \left[ \left( \frac{M}{M_0}\right)^{-\beta} 
		   + \left( \frac{M}{M_0}\right)^{\gamma} 
               \right]^{-1}
  \end{equation}
where $M$ is the hosting DMH mass, $L_{0}$ is a normalization constant, $M_0$
is the critical mass where the luminosity function has a slope change, 
$\beta$ and $\gamma$ are the slopes.
This relation can reproduce the luminosity function with best agreement than the classical Schechter function wich works good for low redshift galaxies.

This equation has a similar fashion to the mass to light relation \citep{vandenbosch03} and the mean relation between stellar mas of a galaxy and the mas of its halo used by \cite{moster10}.