### 2014-07-21 paper/the_bibliography.tex

\begin{thebibliography}{}
\bibitem[Bouwens(2006)]{bouwens06} Bouwens, R. J. et al. 2006, \apj, 653, 53	
		%2006ApJ...653...53B
\bibitem[Bouwens(2012)]{bouwens12} Bouwens, R. J. et al. 2012, \apj, 752, 5 	
		%2012ApJ...752L...5B
\bibitem[Bouwens(2014)]{bouwens14} Bouwens, R. J. et al. 2012, arXiv:1403.4295
\bibitem[Kennicutt(1998)]{kennicutt98} Kennicutt, Robert C., Jr. 1998, \araa, 
36, 189		%1998ARA&A..36..189K
\bibitem[Kennicutt(2009)]{kennicutt09} Kennicutt, Robert C., Jr et al. 2009, \apj, 
703, 4672	%2009ApJ...703.1672K
\bibitem[Law(2011)]{law11} Law, K. et al. 2011, \apj, 738, 124
		%2011ApJ...738..124L
\bibitem[Jiang(2011)]{jiang11} Jiang, Linhua et al. 2011, \apj, 743, 65
		%2011ApJ...743...65J
\bibitem[Lee(2009)]{lee09} Lee, Kyoung-Soo et al. 2009, \apj, 695, 368
		%2009ApJ...695..368L
\bibitem[Lundgren(2014]{lundgren14} Lundgren, Britt F. et al, 2014, \apj, 780,
34		%2014ApJ...780...34L
\bibitem[Madau(1998)]{madau98} Madau, Piero. et al, 1998, \apj, 498, 106M
		%1998ApJ...498..106M
\bibitem[McLure(2009)]{mclure09} McLure, R. J. et al. 2009, \mnras, 395, 2196	
		%2009MNRAS.395.2196M
\bibitem[Moster(2010)]{moster10} Moster, Benjamin P. et al. 2010, \apj, 710, 
903 
\bibitem[Riebe(2013)]{riebe13} Riebe, K. et al. 2013, AN, 334, 691 		
		%2013AN....334..691R
\bibitem[Steidel(2003)]{steidel03} Steidel, Charles C. et al. 2003, \apj, 592, 
728 		%2003ApJ...592..728S
\bibitem[Trimble(1987)]{trimble87} Trible, Virginia. 1987, \araa, 25, 425

\bibitem[van den Bosch(203)]{vandenbosch03} van den Bosh, Frank C. et al. 2003, \mnras,
40, 771		%2003MNRAS.340..771V
\bibitem[Willott(2013)]{willott13} Willott, Chris J. et al. 2013, \aj, 145, 4	
		%2013AJ....145....4W
\end{thebibliography}