### 2015-02-23 paper/methodology.tex

\section{Methodology}

The first goal in this paper is reproduce the observed UVLF from DMH, from
 simulations at redshift $z=6$. The observational datasets where aforementioned.

Bolshoi Simulation \citep{riebe13} was performed by the MuldiDark consortium
MultiDark Database is

UV Luminosity as function of DMH mass. 
Parameters.
Dust Extinction

MCMC, maximum likelihood.

UV luminosity to SFR