### 2015-02-24 paper/discussion.tex
2 paragraphs in methodology

\section{Discussion}

\begin{figure}
\epsscale{1.00}
\plotone{fig/cosmic_variance.pdf}
\caption{Cosmic Variance: The Luminosity Function is made using the DMH catalog 
from the full box and the set of parameters from the small boxes.}
\label{graph_cosmic_variance}
\end{figure}

\citep{lundgren14} SFR evolution from $z=1$ to $6$

\citep{bouwens14} HST Legacy

\citep{jiang11} Keck pectroscopy

%``Fig. 16.— Updated Determinations of the derived SFR (left axis) and U V
%luminosity (right axis) densities versus redshift (5.4). The
%left axis gives the SFR densities we would infer from the measured luminosity
%densities, assuming the Madau et al. (1998) conversion
%factor relevant for star-forming galaxies with ages of 108 yr (see also
%Kennicutt 1998). The right axis gives the U V luminosities we infer
%integrating the present and published LFs to a faint-end limit of −17 mag (0.03
%L∗ )''BOUWENS 2014 UV LF

\begin{figure}
\epsscale{1.00}
\plotone{fig/scattering_plots.png}
\caption{Covariance of parameters for one small box}
\label{graph_scattering_plots}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_wo.pdf}
\caption{Star formation rate as function of the dark matter halo mass without dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region. }
\label{graph_SFR_DMHM_wo}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_w.pdf}
\caption{Star formation rate as function of the dark matter halo mass with dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region. }
\label{graph_SFR_DMHM_w}
\end{figure}