### 2015-02-24 paper/the_bibliography.tex

\begin{thebibliography}{}

\bibitem[Behroozi et al.(2013)]{behroozi13} Behroozi, Peter S., Risa H. Wechsler, 
\and Charlie Conroy. 2013, \apj,  770, 57
% Semi-Analytical Model for SFR-DMHM

\bibitem[Bouwens et al.(2006)]{bouwens06} Bouwens, R. J. et al. 2006, \apj, 653, 53	
		%2006ApJ...653...53B
% Observational luminosity function

\bibitem[Bouwens et al.(2012)]{bouwens12} Bouwens, R. J. et al. 2012, \apj, 752, 5 	
		%2012ApJ...752L...5B
% Observational luminosity function

\bibitem[Bouwens et al.(2014)]{bouwens14} Bouwens, R. J., G. D. Illingworth, P. A. 
Oesch, M. Trenti, I. Labbe', L. Bradley, M. Carollo, et al. 2014, arXiv:1403.4295
% Observational luminosity function

\bibitem[Col\'in et al.(1999)]{colin99}
Col\'n, Pedro, Anatoly A. Klypin, Andrey V. Kravtsov, and Alexei M. Khokhlov. 
1999, \apj , 523,32
% Halo Abundance Matching
\bibitem[Conroy et al.(2006)]{conroy06}
Conroy, Charlie, Risa H. Wechsler, \and Andrey V. Kravtsov. 2006, \apj,647,201
% Halo Abundance Matching

\bibitem[Efstathiou et al.(1988)]{efstathiou88} Efstathiou, G., Richard S. Ellis, 
\and Bruce A. Peterson. 1988, \mnras, 232,431.
% StepWise Maximun Likelihood LF

\bibitem[Finkelstein et al.(2014)]{finkelstein14} Finkelstein, Steven L., Russell E. 
Ryan Jr., Casey Papovich, Mark Dickinson, Mimi Song, Rachel Somerville, 
Henry C. Ferguson, et al. 2014, arXiv:1410.5439
% Observational luminosity function

\bibitem[Kennicutt(1998)]{kennicutt98} Kennicutt, Robert C., Jr. 1998, \araa, 
36, 189		%1998ARA&A..36..189K
\bibitem[Kennicutt et al.(2009)]{kennicutt09} Kennicutt, Robert C., Jr et al. 2009, \apj, 
703, 4672	%2009ApJ...703.1672K

\bibitem[Kravtsov et al.(2004)]{kravtsov04}
Kravtsov, Andrey V., Andreas A. Berlind, Risa H. Wechsler, Anatoly A. Klypin, 
Stefan Gottlöber, Brandon Allgood, \and Joel R. Primack. 2004, \apj, 609, 35
%Halo Abundance Matching

\bibitem[Klypin et al.(2011)]{klypin11} Klypin, Anatoly A., Sebastian Trujillo-Gomez, \
and Joel Primack. 2009, \apj, 740, 102
% Bolshoi 2011

\bibitem[Klypin et al.(2014)]{klypin14} Anatoly Klypin, Gustavo Yepes, Stefan Gottlober, 
Francisco Prada \and Steffen Hess. 2014, arXiv:`1411.4001
% Big Bolshoi Planck 1 - MDPL

\bibitem[Law et al.(2011)]{law11} Law, K. et al. 2011, \apj, 738, 124
		%2011ApJ...738..124L
\bibitem[Jiang et al.(2011)]{jiang11} Jiang, Linhua et al. 2011, \apj, 743, 65
		%2011ApJ...743...65J
\bibitem[Lee et al.(2009)]{lee09} Lee, Kyoung-Soo et al. 2009, \apj, 695, 368
		%2009ApJ...695..368L
\bibitem[Lundgren et al.(2014]{lundgren14} Lundgren, Britt F. et al, 2014, \apj, 780,
34		%2014ApJ...780...34L
\bibitem[Madau et al.(1998)]{madau98} Madau, Piero. et al, 1998, \apj, 498, 106M
		%1998ApJ...498..106M
\bibitem[McLure et al.(2009)]{mclure09} McLure, R. J., M. Cirasuolo, J. S. Dunlop, S. Foucaud, \and O. Almaini. 2009, \mnras, 395, 2196	
		%2009MNRAS.395.2196M
\bibitem[Moster et al.(2010)]{moster10} Moster, Benjamin P. et al. 2010, \apj, 710, 
903

\bibitem[Planck Collaboration et al(2014)]{planck1}
Planck Collaboration, P. A. R. Ade, N. Aghanim, C. Armitage-Caplan, M. Arnaud, 
M. Ashdown, F. Atrio-Barandela, et al. \aap, 571, A16
%Planck 2013 cosmology release.

\bibitem[Prada et al.(2012)]{prada12} Prada, Francisco, Anatoly A. Klypin, 
Antonio J. Cuesta, Juan E. Betancort-Rijo, and Joel Primack. 2012. \mnras, 423, 3018
% Big Bolshoi

\bibitem[Riebe et al.(2013)]{riebe13} Riebe, K. et al. 2013, AN, 334, 691 		
% Multiyear Database Release		%2013AN....334..691R

\bibitem[Schmidt(1968)]{schmidt68} Schmidt, Maarten. 1968, \apj 151, 393.
% StepWise Maximun Likelihood LF

\bibitem[Steidel et al.(2003)]{steidel03} Steidel, Charles C. et al. 2003, \apj, 592, 
728 		%2003ApJ...592..728S
\bibitem[Trimble(1987)]{trimble87} Trible, Virginia. 1987, \araa, 25, 425

\bibitem[van den Bosch et al.(203)]{vandenbosch03} van den Bosh, Frank C. et al. 2003, \mnras,
40, 771		%2003MNRAS.340..771V

\bibitem[Willott et al.(2013)]{willott13} Willott, Chris J., Ross J. McLure, Pascale Hibon,
 Richard Bielby, Henry J. McCracken, Jean-Paul Kneib, Olivier Ilbert, David G. Bonfield,
  Victoria A. Bruce, \and Matt J. Jarvis. 2013, \aj, 145, 4	
		%2013AJ....145....4W
\end{thebibliography}