### 2015-03-18 paper/discussion.tex
Paper: initial results

\section{Discussion}

%\citep{lundgren14} SFR evolution from $z=1$ to $6$

%\citep{bouwens14} HST Legacy

%\citep{jiang11} Keck pectroscopy

%``Fig. 16.— Updated Determinations of the derived SFR (left axis) and U V
%luminosity (right axis) densities versus redshift (5.4). The
%left axis gives the SFR densities we would infer from the measured luminosity
%densities, assuming the Madau et al. (1998) conversion
%factor relevant for star-forming galaxies with ages of 108 yr (see also
%Kennicutt 1998). The right axis gives the U V luminosities we infer
%integrating the present and published LFs to a faint-end limit of −17 mag (0.03
%L∗ )''BOUWENS 2014 UV LF