### 2015-03-18 paper/results.tex

\section{Results}

\begin{figure}
\epsscale{1.00}
\plotone{fig/OD3_w_15.pdf}
\caption{Parameter dispersion: The 10.000 MCMC step parameters histogram for OD3 over the small boxe 15
with dust attenuation.}
\label{fig:ParamDispWithDust}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/OD3_wo_15.pdf}
\caption{Parameter dispersion: The 10.000 MCMC step parameters histogram for OD3 over the small boxe 15
with dust attenuation.}
\label{fig:ParamDispWithoutDust}
\end{figure}

$\gamma$ always has the best coincidence for the minimum at the 10\%.

%\begin{figure}
%\epsscale{1.00}
%\plotone{fig/cosmic_variance.pdf}
%\caption{Cosmic Variance: The Luminosity Function is made using the DMH catalog 
%from the full box and the set of parameters from the small boxes.}
%\label{graph_cosmic_variance}
%\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/bestpar_disp_w_OD3.pdf}
\caption{Cosmic Variance: The best fit parameters histogram for OD3 over the 64 small boxes 
with dust attenuation.}
\label{fig:ParamScatteringWithDust}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/bestpar_disp_w_OD3.pdf}
\caption{Cosmic Variance: The best fit parameters histogram for OD3 over the 64 small boxes 
without dust attenuation.}
\label{fig:ParamScatteringWithoutDust}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/scattering_plots.png}
\caption{Covariance of parameters for one small box}
\label{graph_scattering_plots}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_wo.pdf}
\caption{Star formation rate as function of the dark matter halo mass without dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region. 
Comparison with the GALFORM semi-analitical model \citep{gonzalez14} and a implementation
of abundance matching model \citep{behroozi13}. }
\label{graph_SFR_DMHM_wo}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_w.pdf}
\caption{Star formation rate as function of the dark matter halo mass with dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region.
Comparison with the GALFORM semi-analitical model \citep{gonzalez14} and a implementation
of abundance matching model \citep{behroozi13}. }
\label{graph_SFR_DMHM_w}
\end{figure}